\documentclass[11pt]{article}
\usepackage[left=1.25in,top=1in,right=1.25in,bottom=1.00in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{epsfig}
\usepackage{color}
\usepackage[round]{natbib}
\usepackage{multirow} 
\newcommand{\logit}{\mbox{logit}}
\newcommand{\probit}{\mbox{probit}}
\newcommand{\hiw}{{\small\textsc{HIW}}}
\newcommand{\iw}{{\small\textsc{IW}}}
\newcommand{\N}{\mbox{N}}
\newcommand{\Be}{\mbox{Be}}
\newcommand{\dd}{\mbox{d}}
\newcommand{\C}{\; | \;}
\newcommand{\var}{\text{var}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}




\begin{document}

\title{{\bf Homework Assignment 2}\\Due Friday March 1st midnight}
\author{SDS 384-11 Theoretical Statistics}

\date{}

\maketitle{}
%\textbf{Set algebra and probability laws.}
\begin{enumerate}%\item Read Bertsekas and Tsitsiklis, sections 1.1, 1.2 and 1.3.
%\section{Jeffrey's prior}
%\begin{enumerate}
%\end{enumerate}


\item Remember Hoeffding's Lemma? We proved it with a weaker constant in class using a symmetrization type argument. Now we will prove the original version. Let $X$ be a bounded r.v. in $[a,b]$ such that $E[X]=\mu$.
Let $f(\lambda)=\log E[e^{\lambda(X-\mu)}]$. Show that $f''(\lambda)\leq (b-a)^2/4$. Now use the fundamental theorem of calculus to write $f(\lambda)$ in terms of $f''(\lambda)$ and finish the argument.

\item Bernstein's inequality for bounded i.i.d sequences of random variables $\{X_i\}$ with $|X_i|\leq M$ gives:
$P(|\sum_i (X_i-E[X_i])|\geq t)\leq 2\exp\left(\frac{-t^2/2}{\sum_i \var(X_i)+Mt/3}\right) $. There is another better inequality called Bennett's inequality, which we will prove here. 
\begin{enumerate}
\item Consider zero mean r.v.s $X_i$ such that $|X_i|\leq b$ and $\var(X_i)=\sigma_i^2$. Prove that
$$\log E[\exp(\lambda X_i)]\leq \sigma_i^2\lambda^2\left(\frac{e^{\lambda b}-1-\lambda b}{(\lambda b)^2}\right) \ \ \forall \lambda\in \mathbb{R}.$$
\item Given independent r.v.s $X_i,i=1,\dots,n$ satisfying the above condition prove
$$\mbox{(Bennett's inequality)}\qquad P\left(\sum_i X_i\geq n\delta\right)\leq \exp\left(-\frac{n\sigma^2}{b^2}h(b\delta/\sigma^2)\right),$$
where $\sigma^2=\sum_i\sigma_i^2$ and $h(t):=(1+t)\log (1+t)-t$ for $t\ge 0$.
\item Show that Bennett's inequality is at least as good as Bernstein's inequality.
\end{enumerate}
\item Given a scalar random variable $X$, suppose that there are positive constants $c_1,c_2$ such that, 
$$P(X-E[X]\geq t)\leq c_1\exp(-c_2 t^2)\qquad \forall t\ge 0. $$
\begin{enumerate}
	\item Prove that $\var(X)\leq \frac{c_1}{c_2}$
	\item A median $m_X$ is any number such that $P(X\geq m_X)\geq 1/2$ and $P(X\leq m_X)\geq 1/2$. Show by example that the median does not need to be unique.
	\item Show that if the median concentration bound holds, then for any median $m_X$,
	$$P(|X-m_X|\geq t)\leq c_3\exp(-c_4 t^2),$$
where $c_3=4c_1$ and $c_4=c_2/8$.
	\item Conversely, show that whenever the above median concentration holds, then mean concentration holds with $c_1=2c_3$ and $c_2=c_4/4$.
\end{enumerate}
\item Given a positive semidefinite matrix $Q\in \mathbb{R}^{n\times n}$, consider $Z=\sum_{i,j}Q_{ij}X_iX_j$. When $X_i\sim N(0,1)$, prove the Hanson-Wright inequality.
$$P\left(Z\ge \text{trace}(Q)+ t\right)\leq 2\exp\left(-\min\left\{c_1 t/\|Q\|_{op},c_2 t^2/\|Q\|_F^2\right\}\right),$$
where $\|Q\|_{op}$ and $\|Q\|_{F}$ denote the operator and frobenius norms respectively. \textit{Hint: The rotation-invariance of the
 Gaussian distribution and sub-exponential nature of $\chi^2$-variables could be useful.}
\end{enumerate}
\end{document} 
